Defining Crete's Macroecology based on its soil Microbial Interactome

### 2023-12-24 manuscript.tex
add points with cited references across the document

\author[1,2]{Savvas Paragkamian\ORCID{0000-0002-8508-2521}}
\author[3,4]{Johanna B. Holm\ORCID{0000-0002-7646-4085}}
\author[5]{Christina Pavloudi\ORCID{0000-0001-5106-6067}}
\author[6]{Haris Zafeiropoulos\ORCID{0000-0002-4405-6802}}
\author[1]{Christos Christakis\ORCID{0000-0002-7075-0996}}
\author[1]{Melanthia Stavroulaki\ORCID{0000-0003-4392-7159}}
\author[7]{Stephanie A. Yarwood\ORCID{0000-0003-4593-7908}}
\author[2]{Antonios Magoulas}
\author[1]{Panagiotis F. Sarris\ORCID{0000-0001-7000-8997}}
\author[2]{Giorgos Kotoulas\ORCID{0000-0002-4667-8533}}
\author[3]{Lynn Schriml\ORCID{0000-0001-8910-9851}}
\author[2,$\ast$]{Evangelos Pafilis\ORCID{0000-0001-5079-0125}}

@@

The island of Crete, Greece has been studied since the classical times for its' fauna \citep{Sidiropoulos_Polymeni_Legakis_2017}, flora \citep{Krimbas_2005} and ecosystems \citep{Grove1993}.
Crete is a continental island, fifth largest island of the mediterranean, a mediterranean biodiversity hotspot, model for desertification and climate change, elevation gradient, grasing impacts on ecosystems and evolutionary model.
Yet the terrestrial microbial diversity has been unexplored.

Has been sampled from major worldwide projects of microbiome studies.

@@

Genome Standards Consortium, FAIR data, unrestricted, rich metadata

@@

\subsection{Island Sampling Day: Crete}\label{subsec1}

@@

\section{Equations}\label{sec4}

Equations in \LaTeX{} can either be inline or set as display equations. For
inline equations use the \verb+$...$+ commands. Eg: the equation
$H\psi = E \psi$ is written via the command \verb+$H \psi = E \psi$+.

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum. 
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.

@@

@@

Some Conclusions here.

>>>>

\author[1,2]{Savvas Paragkamian\ORCID{0000-0002-8508-2521}}
\author[3,4]{Johanna B. Holm\ORCID{0000-0002-7646-4085}}
\author[5]{Christina Pavloudi\ORCID{0000-0001-5106-6067}}
\author[6]{Haris Zafeiropoulos\ORCID{0000-0002-4405-6802}}
\author[1]{Christos A. Christakis\ORCID{0000-0002-7075-0996}}
\author[1]{Melanthia Stavroulaki\ORCID{0000-0003-4392-7159}}
\author[7]{Stephanie A. Yarwood\ORCID{0000-0003-4593-7908}}
\author[2]{Antonios Magoulas}
\author[1]{Panagiotis F. Sarris\ORCID{0000-0001-7000-8997}}
\author[2]{Giorgos Kotoulas\ORCID{0000-0002-4667-8533}}
\author[3]{Lynn Schriml\ORCID{0000-0001-8910-9851}}
\author[2,$\ast$]{Evangelos Pafilis\ORCID{0000-0001-5079-0125}}

@@

The island of Crete, Greece has been studied since the classical times for its' fauna \citep{Sidiropoulos_Polymeni_Legakis_2017,Anastasiou2018Tenebrionid}, flora \citep{Krimbas_2005} and ecosystems \citep{Grove1993}.
Crete is a continental island, fifth largest island of the mediterranean, a mediterranean biodiversity hotspot, model for desertification, climate change \citep{Kougioumoutzis2020} and habitat loss \citep{ISPIKOUDIS1993259},
elevation gradient \citep{trigas2013, FAZAN2017}, grasing impacts on ecosystems \citep{JouffroyBapicot2016} and evolutionary model \citep{POULAKAKIS2002}.
Yet the terrestrial microbial diversity has been unexplored.

Has been sampled from major worldwide projects of microbiome studies \citep{Vasar2022, Labouyrie2023, Bahram2018, Orgiazzi2018}. Soil fungi \citep{Mikryukov2023, Davison2021, Tedersoo2021} and other soil eukaryotes \citep{Aslani2022}

@@

Genome Standards Consortium \citep{Field2011}, FAIR data \citep{Wilkinson2016}, unrestricted \citep{Amann2019},
rich metadata and platforms of hosting open data like European Soil Data Centre \citep{Panagos2022}

@@

\subsection{Island Sampling Day: Crete}\label{isd_data}

the data for this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession number PRJEB21776
ENA API \citep{Yuan2023}

\subsection{Data integration}\label{data}

Copernicus CORINE Land Cover \citep{CLC2023}
WorldClim 2.0 \citep{Fick2017}
World Database on Protected Areas (WDPA) \citep{Hanson2022}
Function annotation - FAPROTAX \citep{Louca2016}
Crete geology from IGME
Aridity
%NDVI
Taxonomy Silva 138 \citep{Quast2012}

\subsection{Tools}\label{tools}
dada2 for ASV inferrence \citep{Callahan2016}
PEMA for OTU inferrence \citep{Zafeiropoulos2020}
ape R packege \citep{Paradis2004}
SRS R package \citep{Beule2020}
UMAP ordination \citep{mcinnes2018umap-software}
Numerical ecology analyses with vegan R package \citep{vegan}

Python 3.11.4
Julia Version 1.9.3 \citep{Julia-2017}
R version 4.3.2 \citep{rcoreteam}
Network inferrence using FlashWeave 0.19.2 \citep{Tackmann2019}
Network analysis in igraph R package \citep{Csardi2006}
Differential abundance analysis with ANCOM-BC R package \citep{Lin2020}
Spatial analysis with the sf and stars R packages \citep{Pebesma2023}
U-CIE R package for coloring 3 dimentional data \citep{Koutrouli2022}
ggplot2 \citep{ggplot22016}
pheatmap \citep{Kolde2019}

Computations were performed on HPC infrastrure of HCMR \citep{Zafeiropoulos2021}.

@@

@@

Networks specialists generalists \citep{Barberan2012}

Elevation gradients of biodiversity are known since Humboldt's work \citep{Rahbek2019} 
yet these patterns remain elusive regarding the soil microbiome \citep{Looby2020, Siles2023}.
Mostly because it's difficult to isolate other cofounding effects \citep{Nottingham2018}.

Functions of Richtis gorge are a signal in need to be further investigated.

Enumeratic all biodivarsity \citep{Anthony2023} and decipher the biogeochemical 
processes and interactions of fauna, flora and microbes \citep{Fry2019, Crowther2019,GRANDY201640,Delgado-Baquerizo2020}

Global hotspots \citep{Guerra2022}
and soil conservation instead of specific species \citep{Guerra2021}
and policy \citep{KONINGER2022} across countries \citep{Putten2023}

@@

Mountain peaks soil microbes \citep{Adamczyk2019}
These studies can act as early warning signals towards public health conserns \citep{Banerjee2023}.
Some Conclusions here.
